Which type of analysis to use / using t-test, anova...

- Recognize the differences between one-way, two-way, mixed, or ANCOVA anova test, execute the test and interpret the results
  - Determine if the explanatory variable(s) is continuous, discrete, or mixed
  - Understand the meaning of p-values and how it relates to a normal distribution
- Understand the differences between one-sample t test, two-sample t test, or paired t test, execute the test and interpret the results
  - Learn how to assess an experimental design in order to choose the appropriate t-test
  - Understand probability distributions and how it relates to t-tests

#Loading necessary libraries

```{r}
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
```
#Reading and importing the data

First, we need to look at our data file and understand what we are dealing with. In our example, the penguin data for three different islands is split up between three sheets in the .xlsx file.

We will need to import the correct library that allows us to read .xlsx files in R. Then, we will need to extract each sheet in the .xlsx file into its own dataframe. It would be helpful for us to use the as.data.frame() function to convert the data to dataframes with correct variable types.
```{r}
library(readxl)

torgersenpen <- read_excel("penguins.xlsx", sheet = "Torgersen Island")
biscoepen <- read_excel("penguins.xlsx", sheet = "Biscoe Island")
dreampen <- read_excel("penguins.xlsx", sheet = "Dream Island")

torgersenpen <- as.data.frame(unclass(torgersenpen), stringsAsFactors=TRUE)

torgersenpen |>
  na.omit() -> torgersenpen

biscoepen <- as.data.frame(unclass(biscoepen), stringsAsFactors=TRUE)

biscoepen |> 
  na.omit() ->biscoepen

dreampen <- as.data.frame(unclass(dreampen), stringsAsFactors=TRUE)

dreampen |> 
  na.omit() -> dreampen
```

#Checking dataframe structure using str()
```{r}
str(torgersenpen)
str(biscoepen)
str(dreampen)
```
For the sake of our data exploration, we should have a dataframe which includes all of the data from each separate island. Using the rbind() function, we are able to first bind the Torgersen data with the Biscoe data. Then, we can put it all together using rbind() on the torsergen+biscoe frame with the Dream frame. For safe measure, we will use as.data.frame() on the full set.

```{r}
torgbisc <- rbind(torgersenpen, biscoepen)

full <- rbind(torgbisc, dreampen)
fullframe <- as.data.frame(full)
fullframe
```

Now that our full dataframe is established, we can begin exploring the data. We will first assess whether the bill length is different between the three species. To do this, we will conduct an anova.
  - What type of variable is our explanatory variable? Response variable?
```{r}
fullframe$species <- as.factor(fullframe$species)
billLaov <- aov(bill_length_mm ~ species, data = fullframe)
summary(billLaov)
```

What type of ANOVA did our code conduct? Why was this type of ANOVA conducted? What do the results of our ANOVA tell us?

Next, we will test whether or not the bill length in penguins for Torsergen's Island is sex-dependent. To do this, we will conduct yet another anova to determine the significance (if any) in this relationship.

```{r}
billLaov2 <- aov(bill_length_mm ~ sex, data = torgersenpen)
summary(billLaov2)
```
What type of ANOVA did our code conduct? Why was this type of ANOVA conducted? What do the results of our ANOVA tell us?

I want you to think about the other available statistical tests which are available to us in R, and which would be more appropriate for completing the task of comparing bill lengths between sexes on Torsergen Island. Would the following code be appropriate for the task? Why or why not?

```{r}
t.test(bill_length_mm ~ sex, data = torgersenpen)
```

Finally, we will test the hypothesis that bill length is dependent on the interaction between sex and species. What type of ANOVA would be best for this analysis? Provide the code below.

```{r}
anova3 <- aov(bill_length_mm ~ sex * species, data = fullframe)
summary(anova3)
```

What do the results of this ANOVA tell us? Why are we able to run a two-way ANOVA with the given variables? Are there statistically significant results? Why wouldn't we use a linear model in this type of analysis?


