Exploring normality of data
-determine if a data set is normally distributed or not
-normalize a set of data that isn't normal
-get rid of NA


```{r}
library(readxl)
Torgersen <- read_excel("penguins.xlsx", sheet = "Torgersen Island")
View(Torgersen)
```


```{r}
library(readxl)
Biscoe <- read_excel("penguins.xlsx", sheet = "Biscoe Island")
View(Biscoe)
```

```{r}
library(readxl)
Dream <- read_excel("penguins.xlsx", sheet = "Dream Island")
View(Dream)
```


```{r}
library(readxl)

torgersenpen <- read_excel("penguins.xlsx", sheet = "Torgersen Island")
biscoepen <- read_excel("penguins.xlsx", sheet = "Biscoe Island")
dreampen <- read_excel("penguins.xlsx", sheet = "Dream Island")

torgbisc <- rbind(torgersenpen, biscoepen)

full <- rbind(torgbisc, dreampen)
fullframe <- as.data.frame(full)
```

read in the data
make sure the data is read in correctly
plot the residuals
read the residuals
figure out how to normalize (log)
plot residuals of normalized
get rid of NAs
prep for hypothesis (Bill Length is different between the 3 species, Bill Length in Penguins from Torgersen's Island is Sex Dependent.)

##Check the 
