---
title: "Group Project 1"
subtitle: "Biology 368/664 Bucknell University"
output: github_document
authors: Maria Ruccolo, Ella Boback, Trevor Neff
date: 2 Oct 2024
---

## Start by loading the Data and Packages

Packages are required to run certain funtions in R, the main ones being used here include 'conflicted', 'tidyverse', 'dplyr', 'ggplot2', and 'cowplot'. These same packages can be found in the "Packages" tab in the lower right panel. 

```{r}
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
if (!require("dplyr")) install.packages("dplyr"); library(dplyr)
if (!require("ggplot2")) install.packages("ggplot2"); library(ggplot2)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR) # For the simple.eda function
```


First, we need to look at our data file and understand what we are dealing with. In our example, the penguin data for three different islands is split up between three sheets in the .xlsx file. In order to read in strictly excel files, you must use the read_excel function. It is also important to note that the multiple sheets can cause issues, so you have to read them in sheet by sheet. Binding the data with rbind allows us to have each sheet into one data frame in R. 

We will need to import the correct library that allows us to read .xlsx files in R. Then, we will need to extract each sheet in the .xlsx file into its own dataframe. It would be helpful for us to use the as.data.frame() function to convert the data to dataframes with correct variable types.

```{r}
library(readxl)

torgersenpen <- read_excel("penguins.xlsx", sheet = "Torgersen Island")
biscoepen <- read_excel("penguins.xlsx", sheet = "Biscoe Island")
dreampen <- read_excel("penguins.xlsx", sheet = "Dream Island")

torgbisc <- rbind(torgersenpen, biscoepen)

full <- rbind(torgbisc, dreampen)
fullframe <- as.data.frame(full)
View(fullframe)
```



### Make sure the data is read correctly

Even though you may not be familiar with the data set we are using, we will walk through how to check if your data set was read correctly into R. Checking the rows, columns, and total numbers reassures that no data is missing.

The View() function allows you to see the whole data set completely. 

#Check columns and rows

Looking at these makes sure we have loaded in all the data and each variable. We expect to find 8 columns and 335 rows.

```{r}
nrow(fullframe)
ncol(fullframe)
View(fullframe)
```

#Check top and bottom of data

The top and the bottom also can show us the kind of data we are looking at, as well as the framework of the set.

```{r}
head(fullframe)
tail(fullframe)
```


#Run glimpse and catagorize data

In the bottom left panel, type "?glimpse" and press enter. What does it show you?
Glimpse in the package dplyr allows us to see every column in the data frame. This simply shows you as much data as possible.

Then, we get into the function mutate. This allows us to change the type of variable for our data. R automatically read our data but we need to change the factor variables to factors and the numeric variables to numbers. The as.factor() function is used for the variables sex, island, and species. As you can see, you can also do levels within the factor, as portrayed in the island line of code. 

The next code shows the same process but with the as.numeric function for the numerical values of bill length, bill depth, flipper length, and body mass. 

Using the summary() function, you can see if the as.factor and as.numeric functions worked.

```{r}
glimpse(fullframe)

fullframe |> 
  mutate(sex = as.factor(sex),
         island = factor(island, levels = c("Dream", "Torgersen", "Biscoe")),
         species = as.factor(species)) -> fullframe


fullframe |>
  mutate(bill_length_mm = as.numeric(bill_length_mm),
         bill_depth_mm = as.numeric(bill_depth_mm),
         flipper_length_mm = as.numeric(flipper_length_mm),
         body_mass_g = as.numeric(body_mass_g)) -> fullframe
summary(fullframe)
```

#Check 'n's

Again, although you are likely unfamiliar with this data set, the following code allows you to see the total number of observations made in the data set. If this did not match your original data you collected, you would have been made aware that the data didn't read in correctly. 

```{r}
str(fullframe)
table(fullframe$bill_length_mm)
table(fullframe$sex)
```

### Plot the residuals

First, lets look at the data. Using the ggplot() function, we can create a histogram (geom_histogram()) with the bill length (part of our original hypothesis). It is important to note that each time you create a graph, you have to explicitly state where you want the data pulled from. The theme_cowplot simply makes the graph pretty.

For the second graph, certain features are added (including changing the data frame) to the aesthetic function (aes()) to make the graph specialized. With the colour() fucntion, you can group data within a graph. fill() simply makes it look better (play around with it and see what happends without the fill option!). The binwidth is another way to alter your graph with different widths (play around with this too! Try 0.5, then 3. What looks better?) With the position() function, it just indicates where your legend is. You will learn how to better visualize the data later.

```{r}
ggplot(fullframe) +
  aes(x = bill_length_mm) +
  geom_histogram(bins = 100) +
  theme_cowplot()

ggplot(torgersenpen) +
  aes(x = bill_length_mm, colour = sex, fill = sex) +
  geom_histogram(binwidth = 1.5,  position = "dodge") +
  theme_cowplot()


```

Residuals are the difference between an observed data value and a predicted data value. When we plot them, it allows us to see visually whether or not the data is normal or not. Using lm(), we can look at the residuals of the linear models we make. So, for our first relationship we want to look at (bill length and species), the code follows this pattern:
name of linear model <- lm(bill length depends on(~) species, from the data frame fullframe)

We can do a similar thing with our second relationship (bill length and sex). The one thing that needed to be changed was the fact that the data set torgersenpen had the specific data we needed, without the clutter of the fullframe data set.

plot() graphs the linear models that you just made.

```{r}
lmspecieslength <- lm(bill_length_mm ~ species, data = fullframe)
summary(lmspecieslength)

lmsexlength <- lm(bill_length_mm ~ sex, data = torgersenpen)
summary(lmsexlength)


plot(lmspecieslength)
plot(lmsexlength)
```


### 4. Discuss the residuals

Although these graphs are a lot, the main ones we will focus on (to stay within the scope of this tutorial) are the QQ plots. A QQ plot basically shows how well the data matches the theoretical normal distribution. If all of the points are roughly along the dotted line, that means the data is mostly normally distributed. If the data points do not fit along the line at all it means the data is highly skewed.

For these, it looks like the data is fairly normalized. The data on both the bill length ~ species and the bill length ~ sex look normal with closely fit data on the line. Normally, this would be it for making sure our data is normalized since it already looks normal, but the following code gives some tips on how to normalize a skewed set. 

### 5. Figure out how to normalize the data and plot residuals

Using this following code, this allows us to take log10 of the bill length. Although our data looks fairly normalized, what would happen if you took the log10 of it?


```{r}
fullframe$log10bill_length <- log10(fullframe$bill_length_mm)
simple.eda(fullframe$log10bill_length)
```
As you can see, it doesn't make it that much better, it even makes the QQ plot look worse. Since we are lucky with a normal data set, we don't need to use any normalizing functions. 

#Visualization of Your Data

The Hypotheses that we are testing are:
    -Is Bill Length in Penguins from Torgersen's Island Sex Dependent
    -Is Bill Length different between penguins of different species.
    
We will be visualizing "Is Bill Length in Penguins from Torgersen's Island is Sex Dependent."


First view your data and get familiar with your variables.
In this tutorial we will be testing if the Bill Length is different between the 3 species, and if Bill length in Penguins from Torgersen's Island is Sex Dependent.

There are two simple ways to get a basic view of your data. You can simply type the name of your dataset to view the full data. You can also use summary(YourDataSet) to view the dataset with some summary statistics. Try both bellow!

##View Data
```{r}
summary(torgersenpen)
```

We created a dataset for Penguins in Torgersen's Island called "torgersenpenguins". Simply type in the dataset's name to view the data. Try bellow.

```{r}
torgersenpen
```
It is important to understand the type of variables you have to understand how to best visualize them. There is numerical, or quantitative variables and categorical or qualitative variables. In our data set, type of island, and sex are our categorical variables and bill length is our numerical variable. Therefore, we will be using graphs that will compare categorical and numerical values. These include a Boxplot, Histogram and Bargraph.

To start we will visualize the data comparing Bill Length by Sex to visuzlize if there is a significant differrence in Bill Length, depending on Sex.




##Basic boxplot

Our independent variable is sex, and our dependent variable in this hypothesis is bill length. Therefore we will structure the code boxplot(dependentvariable ~ indapendent variable). xlab and ylab are used to name the x and y axis. main is used to create a title for the whole graph, and col allows us to customize the colors of our graph.

```{r}
boxplot(bill_length_mm ~ sex, data = torgersenpen,
        xlab = "Sex",
        ylab = "Bill Length",
        main = "Boxplot of Bill Length by Sex",
        col = c("lightblue", "lightpink"))

```
##Jitter Boxplot
Now we will add jitter to our boxplot :

The code bellow shows a different way to code for a boxplot using ggplot. In this case, our x(independent) and y(dependent) variables stay the same. We add jitter using the function geom_jitter. We also changed the scale_x_discrete to change the lables of Male and Female on our x axis.

```{r}
ggplot(torgersenpen) +
  aes(x = factor(sex), y = bill_length_mm) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_cowplot() +
  ylab("Bill Length (mm)")+
  xlab(c(" "))+
  scale_x_discrete(labels=c("Female","Male"))

```
###Analyze Boxplot
Now you want to analyze your results and look for any differences in the data for our boxplots :

In both of  boxplot we can see a significant difference between the Bill Length. It seems that males on average have a higher bill length than female penguins from Torgersen's island. Of course we will have to use a statistical test to check if this correlation is actually significant.


##Histogram

Next we will create a Histogram to visualize the same thing in the data. 

In our histogram we want to display the Bill Length on the x-axis and separate these values by female and male.

First lets create a histogram that includes just the Bill Length. To do this simply use the ggplot function as we have before, with the geom_histogram function.
```{r}
ggplot(torgersenpen) +
  aes(x = bill_length_mm) +
  geom_histogram(bins=100) +
  theme_cowplot()+
  xlab("Bill Length (mm)") + 
  ylab("Count")

```

Next lets separate these results by sex : 
To do this, we will use similar code as before, adding aes(x = dependent variable, color = dependent variable, full = dependent variable). This will separate the data by the dependent variable by color.
```{r}
ggplot(torgersenpen) +
  aes(x = bill_length_mm, color = sex, fill = sex) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  theme_cowplot()+
    xlab("Bill Length") + 
  ylab("Count")

```

Finally, lets make these results more aesthetically pleasing :
```{r}
ggplot(torgersenpen) +
  aes(x = bill_length_mm,  color= sex, fill = sex) + 
  geom_density(alpha=.3) +
  theme_cowplot()+
  xlab("Bill Length (mm)") + 
  ylab("Count")
```
###Analyze Histogram
Now you want to analyze your results and look for any differences in the data for our Histogram :

Our Histogram displays similar results to our boxplot, in which we can see a significant difference between the Bill Length seperated by sex. It seems that males on average have a higher bill length than female penguins from Torgersen's island. Of course we will have to use a statistical test to check if this correlation is actually significant.

##Bar Chart
Laslty, we will be making a bar chart using the ggplot function. Our data set and variables stay the same. To create a bar chart we will use the function geom_bar within our code.

```{r}
ggplot(torgersenpen, aes(x = sex, y=bill_length_mm), fill = Type) + 
  geom_bar(stat = "summary", position = "dodge")+ 
  scale_fill_grey()+ 
  labs(title = "Mean Bill Length by Sex (Torgersen Penguins)",
       x = "Sex",
       y = "Mean Bill Length (mm)") +
  theme_cowplot()
```

This will create a very basic Bar Chart to visualize our Data.


###Analyze Bar Chart

As you can see, our graphs are all consistent in that Males appear to have a higher average Bill Lenghth then Female penguins on Torgersen's island. We must test this with statistical models to verify if the correlation is actually there.

##Practice Using Second Hypothesis 
Now we will ask you to test the second hypothesis, using the dataset fullframe, to compare the bill lengths among penguins from different species. There is space for you to practice completing a boxplot, histogram, and barchart. If you are to get stuck, the answers are bellow for each visualization. Also attemtpt to analyze the graphs.


###Practice Boxplot

```{r}

```
Analysis: 

###Practice Histogram

```{r}

```
Analysis: 

###Practice Bar Chart
```{r}

```
Analysis: 

###Answers Boxplot
```{r}
#Without Jitter
boxplot(bill_length_mm ~ species, data = fullframe,
        xlab = "Species",
        ylab = "Bill Length",
        main = "Boxplot of Bill Length by Species",
        col = c("lightblue", "lightpink", "lightgreen"))

#With Jitter
ggplot(fullframe) +
  aes(x = factor(species), y = bill_length_mm) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_cowplot() +
  ylab("Bill Length (mm)")+
  xlab("Species")
```
Analysis: Our boxplot suggests that Chinstrap penguins have the highest average bill length, and adelie have the lowest average bill length.

###Answers Histogram
```{r}
#Just Bill Length
ggplot(fullframe) +
  aes(x = bill_length_mm) +
  geom_histogram(bins=100) +
  theme_cowplot()+
  xlab("Bill Length (mm)") + 
  ylab("Count")

#Bill Length Seperated by Species

ggplot(fullframe) +
  aes(x = bill_length_mm, color = species, fill = species) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  theme_cowplot()+
    xlab("Bill Length") + 
  ylab("Count")

ggplot(fullframe) +
  aes(x = bill_length_mm,  fill = species) + 
  geom_density(alpha=.3) +
  theme_cowplot()+
  xlab("Bill Length (mm)") + 
  ylab("Count")

```
Analysis: Our histogram suggests that Chinstrap penguins have the highest average bill length, and adelie have the lowest average bill length. This is consistent with the boxplot.

###Answers Bar Chart
```{r}
ggplot(fullframe, aes(x = species, y=bill_length_mm), fill = Type) + 
  geom_bar(stat = "summary", position = "dodge")+ 
  scale_fill_grey()+ 
  labs(title = "Mean Bill Length by Island",
       x = "Species",
       y = "Mean Bill Length (mm)") +
  theme_cowplot()
```
Analysis:Our Bar Chart suggests that Chinstrap penguins have the highest average bill length, and adelie have the lowest average bill length. This is consistent with the boxplots and the histograms.


##Analyzing Data using t-test, and Anova
Now you will learn how to analyze data using T-tests and Anova.

First, Which type of analysis to use / using t-test, anova...

- Recognize the differences between one-way, two-way, mixed, or ANCOVA anova test, execute the test and interpret the results
  - Determine if the explanatory variable(s) is continuous, discrete, or mixed
  - Understand the meaning of p-values and how it relates to a normal distribution
- Understand the differences between one-sample t test, two-sample t test, or paired t test, execute the test and interpret the results
  - Learn how to assess an experimental design in order to choose the appropriate t-test
  - Understand probability distributions and how it relates to t-tests

#Checking dataframe structure using str()
```{r}
str(torgersenpen)
str(biscoepen)
str(dreampen)
```

Since our full dataframe is already established, we can begin exploring the data. We will first assess whether the bill length is different between the three species. To do this, we will conduct an anova.
  - What type of variable is our explanatory variable? Response variable?

```{r}
billLaov <- aov(bill_length_mm ~ species, data = fullframe)
summary(billLaov)
```

What type of ANOVA did our code conduct? Why was this type of ANOVA conducted? What do the results of our ANOVA tell us?

Next, we will test whether or not the bill length in penguins for Torsergen's Island is sex-dependent. To do this, we will conduct yet another anova to determine the significance (if any) in this relationship.

```{r}
billLaov2 <- aov(bill_length_mm ~ sex, data = torgersenpen)
summary(billLaov2)
```

What type of ANOVA did our code conduct? Why was this type of ANOVA conducted? What do the results of our ANOVA tell us?

I want you to think about the other available statistical tests which are available to us in R, and which would be more appropriate for completing the task of comparing bill lengths between sexes on Torsergen Island. Would the following code be appropriate for the task? Why or why not?

```{r}
t.test(bill_length_mm ~ sex, data = torgersenpen)
```

Finally, we will test the hypothesis that bill length is dependent on the interaction between sex and species. What type of ANOVA would be best for this analysis? Provide the code below.

```{r}
anova3 <- aov(bill_length_mm ~ sex * species, data = fullframe)
summary(anova3)
```

What do the results of this ANOVA tell us? Why are we able to run a two-way ANOVA with the given variables? Are there statistically significant results? Why wouldn't we use a linear model in this type of analysis?






