---
title: "Group Project 1"
subtitle: "Biology 368/664 Bucknell University"
output: pdf_document
authors: Maria Ruccolo, Ella Boback, Trevor Neff
date: 2 Oct 2024
---

### Start by loading the data and packages

First, we need to look at our data file and understand what we are dealing with. In our example, the penguin data for three different islands is split up between three sheets in the .xlsx file. In order to read in strictly excel files, you must use the read_excel function. It is also important to note that the multiple sheets can cause issues, so you have to read them in sheet by sheet. Binding the data with rbind allows us to have each sheet into one data frame in R. 

We will need to import the correct library that allows us to read .xlsx files in R. Then, we will need to extract each sheet in the .xlsx file into its own dataframe. It would be helpful for us to use the as.data.frame() function to convert the data to dataframes with correct variable types.

```{r}
library(readxl)

torgersenpen <- read_excel("penguins.xlsx", sheet = "Torgersen Island")
biscoepen <- read_excel("penguins.xlsx", sheet = "Biscoe Island")
dreampen <- read_excel("penguins.xlsx", sheet = "Dream Island")

torgbisc <- rbind(torgersenpen, biscoepen)

full <- rbind(torgbisc, dreampen)
fullframe <- as.data.frame(full)
View(fullframe)
```

#load required packages

Packages are required to run certain funtions in R, the main ones being used here include 'conflicted', 'tidyverse', 'dplyr', 'ggplot2', and 'cowplot'. These same packages can be found in the "Packages" tab in the lower right panel. 

```{r}
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
if (!require("dplyr")) install.packages("dplyr"); library(dplyr)
if (!require("ggplot2")) install.packages("ggplot2"); library(ggplot2)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR) # For the simple.eda function
```


### Make sure the data is read correctly

Even though you may not be familiar with the data set we are using, we will walk through how to check if your data set was read correctly into R. Checking the rows, columns, and total numbers reassures that no data is missing.

The View() function allows you to see the whole data set completely. 

#Check columns and rows

Looking at these makes sure we have loaded in all the data and each variable.

```{r}
nrow(fullframe)
ncol(fullframe)
View(fullframe)
```

#Check top and bottom of data

The top and the bottom also can show us the kind of data we are looking at, as well as the framework of the set.

```{r}
head(fullframe)
tail(fullframe)
```


#Run glimpse and catagorize data

In the bottom left panel, type "?glimpse" and press enter. What does it show you?
Glimpse in the package dplyr allows us to see every column in the data frame. This simply shows you as much data as possible.

Then, we get into the function mutate. This allows us to change the type of variable for our data. R automatically read our data but we need to change the factor variables to factors and the numeric variables to numbers. The as.factor() function is used for the variables sex, island, and species. As you can see, you can also do levels within the factor, as portrayed in the island line of code. 

The next code shows the same process but with the as.numeric function for the numerical values of bill length, bill depth, flipper length, and body mass. 

Using the summary() function, you can see if the as.factor and as.numeric functions worked.

```{r}
glimpse(fullframe)

fullframe |> 
  mutate(sex = as.factor(sex),
         island = factor(island, levels = c("Dream", "Torgersen", "Biscoe")),
         species = as.factor(species)) -> fullframe


fullframe |>
  mutate(bill_length_mm = as.numeric(bill_length_mm),
         bill_depth_mm = as.numeric(bill_depth_mm),
         flipper_length_mm = as.numeric(flipper_length_mm),
         body_mass_g = as.numeric(body_mass_g)) -> fullframe
summary(fullframe)
```

#Check 'n's

Again, although you are likely unfamiliar with this data set, the following code allows you to see the total number of observations made in the data set. If this did not match your original data you collected, you would have been made aware that the data didn't read in correctly. 

```{r}
str(fullframe)
table(fullframe$bill_length_mm)
table(fullframe$sex)
```

### Plot the residuals

First, lets look at the data. Using the ggplot() function, we can create a histogram (geom_histogram()) with the bill length (part of our original hypothesis). It is important to note that each time you create a graph, you have to explicitly state where you want the data pulled from. The theme_cowplot simply makes the graph pretty.

For the second graph, certain features are added (including changing the data frame) to the aesthetic function (aes()) to make the graph specialized. With the colour() fucntion, you can group data within a graph. fill() simply makes it look better (play around with it and see what happends without the fill option!). The binwidth is another way to alter your graph with different widths (play around with this too! Try 0.5, then 3. What looks better?) With the position() function, it just indicates where your legend is. You will learn how to better visualize the data later.

```{r}
ggplot(fullframe) +
  aes(x = bill_length_mm) +
  geom_histogram(bins = 100) +
  theme_cowplot()

ggplot(torgersenpen) +
  aes(x = bill_length_mm, colour = sex, fill = sex) +
  geom_histogram(binwidth = 1.5,  position = "dodge") +
  theme_cowplot()


```

Residuals are the difference between an observed data value and a predicted data value. When we plot them, it allows us to see visually whether or not the data is normal or not. Using lm(), we can look at the residuals of the linear models we make. So, for our first relationship we want to look at (bill length and species), the code follows this pattern:
name of linear model <- lm(bill length depends on(~) species, from the data frame fullframe)

We can do a similar thing with our second relationship (bill length and sex). The one thing that needed to be changed was the fact that the data set torgersenpen had the specific data we needed, without the clutter of the fullframe data set.

plot() graphs the linear models that you just made.

```{r}
lmspecieslength <- lm(bill_length_mm ~ species, data = fullframe)
summary(lmspecieslength)

lmsexlength <- lm(bill_length_mm ~ sex, data = torgersenpen)
summary(lmsexlength)


plot(lmspecieslength)
plot(lmsexlength)
```


### 4. Discuss the residuals

Although these graphs are a lot, the main ones we will focus on (to stay within the scope of this tutorial) are the QQ plots. A QQ plot basically shows how well the data matches the theoretical normal distribution. If all of the points are roughly along the dotted line, that means the data is mostly normally distributed. If the data points do not fit along the line at all it means the data is highly skewed.

For these, it looks like the data is fairly normalized. The data on both the bill length ~ species and the bill length ~ sex look normal with closely fit data on the line. Normally, this would be it for making sure our data is normalized since it already looks normal, but the following code gives some tips on how to normalize a skewed set. 

### 5. Figure out how to normalize the data and plot residuals

Using this following code, this allows us to take log10 of the bill length. Although our data looks fairly normalized, what would happen if you took the log10 of it?


```{r}
fullframe$log10bill_length <- log10(fullframe$bill_length_mm)
simple.eda(fullframe$log10bill_length)
```
As you can see, it doesn't make it that much better, it even makes the QQ plot look worse. Since we are lucky with a normal data set, we don't need to use any normalizing functions. 

##Visualization of Your Data

The Hypothesis that we are testing are:
    -Is Bill Length in Penguins from Torgersen's Island is Sex Dependent
    -Is Bill Length is different between the 3 species
    
We will be visualizing "Is Bill Length in Penguins from Torgersen's Island is Sex Dependent."


First view your data and get familiar with your variables.
In this tutorial we will be testing if the Bill Length is different between the 3 species, and if Bill length in Penguins from Torgersen's Island is Sex Dependent.

There are two simple ways to get a basic view of your data. You can simply type the name of your dataset to view the full data. You can also use summary(YourDataSet) to view the dataset with some summary statistics. Try both bellow!

#View Data
```{r}
summary(torgersenpen)
```

We created a dataset for Penguins in Torgersen's Island called "torgersenpenguins". Simply type in the dataset's name to view the data. Try bellow.

```{r}
torgersenpen
```

Next we will visualize the data comparing Bill Length by Sex to visuzlize if there is a significant differrence in Bill Length, depending on Sex.

# Basic boxplot
```{r}
boxplot(bill_length_mm ~ sex, data = torgersenpen,
        xlab = "Sex",
        ylab = "Bill Length",
        main = "Boxplot of Bill Length by Sex",
        col = c("lightblue", "lightpink"))

```
Now we will add jitter to our boxplot :

```{r}
ggplot(torgersenpen) +
  aes(x = factor(sex), y = bill_length_mm) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_cowplot() +
  ylab("Bill Length (mm)")+
  xlab(c(" "))+
  scale_x_discrete(labels=c("Female","Male"))

```
##Analyze Graph
Now you want to analyze your results and look for any differences in the data for our boxplots :

In both of  boxplot we can see a significant difference between the Bill Length. It seems that males on average have a higher bill length than female penguins from Torgersen's island. Of course we will have to use a statistical test to check if this correlation is actually significant.


#Histogram

Next we will create a Histogram to visualize the same thing in the data. 

In our histogram we want to display the Bill Length on the x-axis and separate these values by female and male.

First lets create a histogram that includes just the Bill Lenght.
```{r}
ggplot(torgersenpen) +
  aes(x = bill_length_mm) +
  geom_histogram(bins=100) +
  theme_cowplot()+
  xlab("Bill Length (mm)") + 
  ylab("Count")

```

Next lets separate these results by sex :
```{r}
ggplot(torgersenpen) +
  aes(x = bill_length_mm, color = sex, fill = sex) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  theme_cowplot()+
    xlab("Bill Length") + 
  ylab("Count")

```
Finally, lets make these results look better in our graph :
```{r}
ggplot(torgersenpen) +
  aes(x = bill_length_mm,  fill = sex) + 
  geom_density(alpha=.3) +
  theme_cowplot()+
  xlab("Bill Length (mm)") + 
  ylab("Count")
```
##Analyze Graph
Now you want to analyze your results and look for any differences in the data for our Histogram :

Our Histogram displays similar results to our boxplot, in which we can see a significant difference between the Bill Length seperated by sex. It seems that males on average have a higher bill length than female penguins from Torgersen's island. Of course we will have to use a statistical test to check if this correlation is actually significant.

#Bar Chart
```{r}
# Calculate means for bill length by sex
summary_data <- torgersenpen %>%
  group_by(sex) %>%
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))

# Create bar plot without error bars
ggplot(summary_data, aes(x = sex, y = mean_bill_length)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Mean Bill Length by Sex (Torgersen Penguins)",
       x = "Sex",
       y = "Mean Bill Length (mm)") +
  theme_cowplot()
```

This will create a very basic Bar Chart to visualize our Data.

Next we will add error lines to this barchart:

```{r}
# Calculate means and standard errors for bill length by sex
summary_data <- torgersenpen %>%
  group_by(sex) %>%
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
            se_bill_length = sd(bill_length_mm, na.rm = TRUE) / sqrt(n()))

# Create bar plot with error bars
ggplot(summary_data, aes(x = sex, y = mean_bill_length)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_errorbar(aes(ymin = mean_bill_length - se_bill_length,
                    ymax = mean_bill_length + se_bill_length), 
                width = 0.2) +
  labs(title = "Bill Length by Sex (Torgersen Penguins)",
       x = "Sex",
       y = "Mean Bill Length (mm)") +
  theme_cowplot()
```
##Analyze Bar Charts

As you can see, our graphs are all consistent in that Males appear to have a higher average Bill Lenghth then Female penguins on Torgersen's island. We must test this with statistical models to verify if the correlation is actually there.


##Analyzing Data using t-test, and Anova
Now you will learn how to analyze data using T-tests and Anova.

First, Which type of analysis to use / using t-test, anova...

- Recognize the differences between one-way, two-way, mixed, or ANCOVA anova test, execute the test and interpret the results
  - Determine if the explanatory variable(s) is continuous, discrete, or mixed
  - Understand the meaning of p-values and how it relates to a normal distribution
- Understand the differences between one-sample t test, two-sample t test, or paired t test, execute the test and interpret the results
  - Learn how to assess an experimental design in order to choose the appropriate t-test
  - Understand probability distributions and how it relates to t-tests

#Checking dataframe structure using str()
```{r}
str(torgersenpen)
str(biscoepen)
str(dreampen)
```
For the sake of our data exploration, we should have a dataframe which includes all of the data from each separate island. Using the rbind() function, we are able to first bind the Torgersen data with the Biscoe data. Then, we can put it all together using rbind() on the torsergen+biscoe frame with the Dream frame. For safe measure, we will use as.data.frame() on the full set.

```{r}
torgbisc <- rbind(torgersenpen, biscoepen)

full <- rbind(torgbisc, dreampen)
fullframe <- as.data.frame(full)
fullframe
```

Now that our full dataframe is established, we can begin exploring the data. We will first assess whether the bill length is different between the three species. To do this, we will conduct an anova.
  - What type of variable is our explanatory variable? Response variable?

```{r}
fullframe$species <- as.factor(fullframe$species)
billLaov <- aov(bill_length_mm ~ species, data = fullframe)
summary(billLaov)
```

What type of ANOVA did our code conduct? Why was this type of ANOVA conducted? What do the results of our ANOVA tell us?

Next, we will test whether or not the bill length in penguins for Torsergen's Island is sex-dependent. To do this, we will conduct yet another anova to determine the significance (if any) in this relationship.

```{r}
billLaov2 <- aov(bill_length_mm ~ sex, data = torgersenpen)
summary(billLaov2)
```

What type of ANOVA did our code conduct? Why was this type of ANOVA conducted? What do the results of our ANOVA tell us?

I want you to think about the other available statistical tests which are available to us in R, and which would be more appropriate for completing the task of comparing bill lengths between sexes on Torsergen Island. Would the following code be appropriate for the task? Why or why not?

```{r}
t.test(bill_length_mm ~ sex, data = torgersenpen)
```

Finally, we will test the hypothesis that bill length is dependent on the interaction between sex and species. What type of ANOVA would be best for this analysis? Provide the code below.

```{r}
anova3 <- aov(bill_length_mm ~ sex * species, data = fullframe)
summary(anova3)
```

What do the results of this ANOVA tell us? Why are we able to run a two-way ANOVA with the given variables? Are there statistically significant results? Why wouldn't we use a linear model in this type of analysis?






