---
title: "Normality 2"
output: html_document
date: "2024-09-30"
---

#Load In Data
```{r}
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
```

#Load in Excell File
```{r}
library(readxl)

torgersenpen <- read_excel("penguins.xlsx", sheet = "Torgersen Island")

torgersenpen <- as.data.frame(unclass(torgersenpen), stringsAsFactors=TRUE)

torgersenpen |>
  drop_na(sex, bill_length_mm) -> torgersenpenguins
```


#View Data
```{r}
summary(torgersenpenguins)
```

##Hypothesis that we are going to visualize are:
    -Is Bill Length is different between the 3 species
    -Is Bill Length in Penguins from Torgersen's Island is Sex Dependent


First view your data and get familiar with your variables.
In this tutorial we will be testing if the Bill Length is different between the 3 species, and if Bill length in Penguins from Torgersen's Island is Sex Dependent.


We created a dataset for Penguins in Torgersen's Island called "torgersenpenguins". Simply type in the dataset's name to view the data. Try bellow.

```{r}
torgersenpenguins
```

# Basic boxplot
```{r}
boxplot(bill_length_mm ~ sex, data = torgersenpenguins,
        xlab = "Sex",
        ylab = "Bill Length",
        main = "Boxplot of Bill Length by Sex",
        col = c("lightblue", "lightpink"))

```

#Histogram

```{r}
ggplot(torgersenpenguins) +
  aes(x = bill_length_mm) +
  geom_histogram(bins=100) +
  theme_cowplot()+
  xlab("Egg Mass") + 
  ylab("Count")

ggplot(torgersenpen) +
  aes(x = bill_length_mm, color = sex, fill = sex) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  theme_cowplot()+
    xlab("Bill_length") + 
  ylab("Count")

ggplot(torgersenpen) +
  aes(x = bill_length_mm,  fill = sex) + 
  geom_density(alpha=.3) +
  theme_cowplot()+
  xlab("Bill Length") + 
  ylab("Count")
```

