Visualizing Data 
- load data in -> make different columns for each island
-make specific types of graph / edit them (Specific goals TBD)
-hypothesis : Bill Length is different between the 3 species
              Bill Length in Penguins from Torgersen's Island is Sex Dependent.
-box-plot
-scatter-plot




```{r}
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted) # For dealing with conflicts
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # For everything
conflict_prefer_all("dplyr", quiet = TRUE)
```


```{r}
library(readxl)

torgersenpen <- read_excel("penguins.xlsx", sheet = "Torgersen Island")

torgersenpen <- as.data.frame(unclass(torgersenpen), stringsAsFactors=TRUE)

torgersenpen |>
  drop_na() -> Completetorgersenpen

```
```

The hypothesis that we are going to visualize are:
    -Is Bill Length is different between the 3 species
    -Is Bill Length in Penguins from Torgersen's Island is Sex Dependent


##BoxPlot with Jitter

First view your data and get familiar with your variables.
In this tutorial we will be testing if the Bill Length is different between the 3 species, and if Bill length in Penguins from Torgersen's Island is Sex Dependent.


We created a dataset for Penguins in Torgersen's Island called "torgersenpen". Simply type in the dataset's name to view the data. Try bellow.

```{r}
summary(torgersenpen)
```

First we will visualize if Bill Length in Penguins from Torgersen's Island is Sex Dependent. To do this we will create a boxplot looking at Bill Length of Penguins from Torgersen's Island seperated by Sex.

```{r}
ggplot(torgersenpen) +
  aes(x = factor(sex), y = bill_length_mm) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(color="black",) +
  theme_cowplot()
```

##Histogram

```{r}
ggplot(torgersenpen) +
  aes(x = bill_length_mm) +
  geom_histogram(bins=100) +
  theme_cowplot()+
  xlab("Egg Mass") + 
  ylab("Count")

ggplot(torgersenpen) +
  aes(x = bill_length_mm, color = sex, fill = sex) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  theme_cowplot()+
    xlab("Bill_length") + 
  ylab("Count")

ggplot(torgersenpen) +
  aes(x = bill_length_mm,  fill = sex) + 
  geom_density(alpha=.3) +
  theme_cowplot()+
  xlab("Bill Length") + 
  ylab("Count")
```

